A Plasma Torus Around a Young Low-Mass Star

### 2025-05-28 paper/ms.tex
update

Stars with masses below about half that of the Sun, M dwarfs, are the
only type of star to offer near-term prospects for detecting the
atmospheres of rocky exoplanets with surface water.  Community
investment with JWST is proceeding accordingly
\citep[][]{Redfield2024,TRAPPIST1JWSTCommunityInitiative2024}.  The
question of how M dwarfs influence their planets---especially the
retention of their atmospheres---has correspondingly grown in
importance.  Previous work has established
that most M dwarfs host close-in planets \citep{Dressing2015} that on
average are subject to long circumstellar disk lifetimes
\citep{Ribas2015}, to high doses of UV radiation \citep{France2016},
and to a high incidence of flares 
\citep{Feinstein2020}.  However, despite extensive work in these
areas, the plasma and magnetospheric environments that bathe young,
close-in exoplanets remain challenging to quantify.  Understanding
these environments is crucial because they directly impact atmospheric
retention and habitability of close-in exoplanets.

@@

The two leading hypotheses for explaining CPVs are either that
transiting clumps of circumstellar material corotate with the star
\citep{Stauffer2017,Gunther2022,Bouma2024}, or that these stars
represent an extreme in naturally-occurring distributions of starspots
or faculae \citep{Koen2021}.  The main argument against a
starspot-only explanation invokes the timescales and amplitudes of the
sharpest photometric features.  However, no independent evidence has
yet been acquired for the presence of any circumstellar material.  If
such material exists, then the geometric correction from the transit
probability would imply an intrinsic occurrence rate at least a few
times larger than the observed rate, suggesting that these clumps
could exist around $\gtrsim$10\% of M dwarfs during their early lives.

@@

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.925\textwidth]{f2.pdf}
  \vspace{-0.2cm}
  \caption{{\bf Evolution of TIC\,141146667 during the Keck/HIRES
  observation (green bar).}  {\bf a,} TESS simple aperture photometry.
  Data gaps caused by light from the Earth (BTJD 3356-3358.5) and Moon
  (BTJD 3359.5-3360.5) are visible.  Raw two minute
  data are in gray; black time-averages to ten minute sampling.  {\bf
  b-c,} Folded TESS light curve before and after spectroscopy.  Black
  now phase-averages to 100 points per
	3.93\ hour cycle.  During BTJD 3352-3356, a state switch occurred near BTJD
  3353, and the dip at $\phi$$\approx$0.8 disappeared.  While the
  large flux decrement was present both before and after the HIRES
  sequence, its photometric shape evolved during the data gap.  The
  orange bar denotes times of spectroscopic transits for the inner two
  H$\alpha$ clumps observed with HIRES (see Figure~\ref{fig:spec}).}
  \label{fig:fulllc}
\end{figure*}

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.99\textwidth]{f3.pdf}
  \vspace{-0.25cm}
  \captionsetup{labelformat=moviefmt,labelsep=colon}
	\caption{\textbf{Emission from circumstellar plasma orbiting
TIC\,141146667.}
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{{\bf
  available here}},
  shows the spectral evolution over five hours.
  {\bf a,} Average TESS light curve from 5 February 2024 to 26
  February 2024 folded on the \periodhr\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf b,} Keck/HIRES H$\alpha$ spectra acquired on 17
  February 2024.  The continuum is set to unity, and the darkest color
  is at twice the continuum to accentuate emission outside the line
  core ($|v/v_{\rm eq}|>1$, for $v_{\rm eq}$=130\,\kms).  While
  emission in the line core originates in the star's chromosphere,
  the sinusoidal emission features are most readily described by a
  warped plasma torus.
  {\bf c,} Individual epochs of Panel b, visible in the
  online movie.  The dotted line shows the time-averaged spectrum,
  $f_{\langle t \rangle}$.
  {\bf d,} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf e,} As in Panel b, after subtracting the time-averaged
  spectrum.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch
  is set to mirror the dynamic range of the data.
  {\bf f,} Individual epochs of Panel e, visible in the online
  movie.}
  \label{fig:spec}
\end{figure*}

@@

We observed TIC\,141146667 ($V$=16.2) for five hours on
17~February~2024 using the High Resolution Echelle Spectrometer
(HIRES; \citealt{vogt_hires_1994}) on the 10\,m Keck I telescope.  The
observations spanned the second half of the night, from 17 February
2024 10:47 to 16:13 (UT).  The star's airmass spanned $z$=1.2-2.2, and
we opted for a fixed 15 minute cadence, except for a final 10 minute
exposure due to increasing sky brightness at sunrise.  We observed
without the iodine cell and used the C2 decker
(0$\farcs$86$\times$14$\farcs$0) in the red configuration, yielding a
spectral resolution $R$$\approx$45{,}000 ($\delta
v$$\approx$6.7\,\kms).  We
binned the CCD readout by a factor of three in the spatial dimension,
yielding $\approx$1,000 photons (S/N=33) per pixel in the continuum at
6500\,\AA.  Strong winds contributed to
1$''$-1\farcs5 seeing over the night, but conditions were otherwise
favorable.  We reduced the echelleogram to a one-dimensional spectrum
using the standard techniques of the California Planet Survey
\citep{Howard2010}.  
%Figure~\ref{fig:spec} shows the result in the
%vicinity of H$\alpha$ without any additional processing.

TESS observed TIC\,141146667 ($T$=13.3) for six non-contiguous months
spanning 2019-2024.  TESS acquired these observations at two-minute
cadence during Sectors 41, 48 (TESS DDT039, PI: Kunimoto), and 75
(TESS Program G06030, PI: Bouma).   The 30-minute data during Sectors
14, 15, and 21 smeared sharp features over the star's 3.93\,hour
period (see \citealt{Gunther2022}).  The movie in
Figure~\ref{fig:lc} shows the two-minute data: 
like other CPVs, TIC\,141146667 maintains a fixed period over
timescales of years while its detailed photometric morphology evolves.
The nearest known star, TIC\,141146666 ($T$=14.5), is 25$''$ from
TIC\,141146667 and is photometrically stable in the TESS images;
crowding is not a concern.

We observed with Keck during Sector~75 in an attempt to obtain
simultaneous observations.   Figure~\ref{fig:fulllc} shows the result;
Earth passed within 25$^\circ$ of the TESS camera's boresight from BTJD
3356.0-3358.5, which caused a data gap that ended twelve hours after our
Keck/HIRES observations (green bar).  From BTJD 3359.4-3362.0, the Moon
then passed within 25$^\circ$ of the camera's boresight.  Based on the
level of scattered light in the optimal aperture
\citep{Jenkins2016}, we manually masked out data in the TESS light curve
from 3359.40-3360.13; the remainder of the data during the lunar
approach were usable.  Small gaps from BTJD 3353.55-3353.77 and
3360.12-3360.33 were caused by data downlinks at the spacecraft's
perigee and apogee.

Although the data gap is unfortunate, Figure~\ref{fig:fulllc} shows that
both before and after the HIRES data were acquired, a large flux
decrement spanned roughly half of each cycle.  From BTJD 3352-3356, this
dip had two sharp local minima;  the minimum at $\phi$$\approx$0.8
became less pronounced following the flare at BTJD 3353, yielding a dip more
closely resembling an asymmetric ``V'' than a ``W''.  Similar CPV state
changes have been previously noted \citep{Stauffer2017,Bouma2024}.  The
photometric shape therefore evolved during the twelve cycles spanning
the 3356-3358 gap, since the average shape from 3358-3363 more closely
resembles the initial ``W''.  Nonetheless, the general photometric
morphology---a small brightening over 45\% of the period, followed by a
complex flux dip spanning 55\% of the period---is similar before and
after the data gap.

@@

Figure~\ref{fig:spec} compiles the TESS and HIRES data from February
2024.  The HIRES spectra \added{(Panels b,c,e,f)} show emission in H$\alpha$ beyond the star's
equatorial velocity, $v_{\rm eq}$, of 130\,\kms.  There are at least
two distinct emission components, separated by 180$^\circ$ in phase.
The inner component at lower velocities has clearer sinusoidal
behaviour in time and is double-peaked, with peak semi-amplitudes of
2.07\,$v_{\rm eq}$ and 2.88\,$v_{\rm eq}$, measured following the
procedure in Appendix~\ref{subsec:halpha}.  There is significant
non-periodic variability in the emissivity of this double-peaked
component: the flux excess begins with an amplitude 70\% that of the
continuum, and diminishes to 10\% by the end.  The higher-velocity
component 180$^\circ$ opposite in phase is detected only from
$\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this outer component appears
connected to the star in velocity space.  While its peak
semi-amplitude of 3.88\,$v_{\rm eq}$ is achieved at both $\phi$=0.25
and 0.75, its amplitude similarly decreases from a 60\% excess over
the continuum at the beginning of the observation sequence to a 10\%
excess by its end.  
From the same procedure in Appendix~\ref{subsec:halpha},
we measured
the period for all three spectroscopic
emission components to be consistent with the photometric \periodhr\ hour
period, to within two minutes for the lower-velocity component.

These sinusoidal emission features require circumstellar clumps of
partially-ionized hydrogen to corotate with the star.  Based on the
observed sinusoidal periods and velocities, this material's motion is
not controlled by gravitational attraction to the star; it is 
a plasma being dragged along with a rigidly rotating stellar magnetic
field.  The velocity semi-amplitude of the sinusoids gives the mean
distance of each clump from the stellar \added{rotation axis}: 2.07\,$R_\star$ and
2.88\,$R_\star$ for the inner clumps, and 3.88\,$R_\star$ for the outer
clump.   These clumps transit in front of the star when passing from
negative to positive velocity.  The transits of the two inner clumps
last $\approx$22\% of each cycle, from $\phi$=-0.1 to $\phi$=+0.12 (see
Appendix~\ref{subsec:halpha}).  This spectroscopic transit coincides
with the latter half of the complex eclipse feature in the TESS data
\added{(see Figure~\ref{fig:fulllc} and Panels a and b of Figure~\ref{fig:spec})}.

The H$\alpha$ line core\added{, visible in Panels e and f of
Figure~\ref{fig:spec}, }is more complex.  At $|\Delta v / v_{\rm
eq}|<1$, most observed H$\alpha$ photons come from the star's
chromosphere; circumstellar material might also modulate the line
profile.  Figure~\ref{fig:spec}e suggests line core variability caused
by both bright and dark regions on the star's surface, superposed with
smaller-amplitude variability from the transiting circumstellar
material.  For instance, from $\phi$=0-0.3, the double-peaked emission
feature is visible when viewed both on and off-limb; this feature is
circumstellar in origin.  However, the large emission feature that
crosses the star from $\phi$=0.4-0.9 emits at an amplitude greater
than that observed from the circumstellar components, and it crosses
the stellar velocity surface at a speed that suggests it instead
originates from a chromospherically bright region on the star's
surface.  Similarly, from $\phi$=0.6-1.15 a 20\% deep absorption
feature slowly crosses the H$\alpha$ line profile.  This feature
suggests either a chromospherically dark region (e.g.~a starspot
group) crossing the star's surface, or an azimuthally extended
absorptive component to the circumstellar material which is not
visible in emission.  The origin of the other bright and dark
streaks passing across the line core \added{in Figure~\ref{fig:spec}e
}are similarly ambiguous.  \added{Figure~\ref{fig:spec}d shows} a
final exercise to quantify the behavior of the line core, where
$f_{\rm H\alpha\ core}$ denotes the summed \added{H$\alpha$} flux at
$|\Delta v / v_{\rm eq}|<1$.  This panel shows that changes in the
line core flux correlate with the broadband variability throughout
most of the light curve, except near $\phi$$\approx$0.5, corresponding
to the transit of the 3.9\,$R_\star$ clump and the occultation of the
lower-velocity clump.  \added{Spectroscopic transits therefore do not
necessarily yield photometric transits.}

@@

While the H$\alpha$ emission seems to mostly originate from clumps, the
complex flux dip in TIC\,141146667 spans roughly half of each cycle; in
CPVs more broadly, individual dips can last 5-50\% of each cycle, and
they are often distributed in longitude \citep{Bouma2024}.  This raises
the possibility that azimuthally-distributed material may be the norm
for CPVs, and that a warped torus may be a more accurate picture than a
clump.

\citet{Townsend2005} outlined the physics of how rapidly-rotating
stars with strong magnetic fields can support plasma tori \added{\citep[see
also][]{Nakajima1985,Ferreira2000,Petit2013,Daley-Yates2024}}.  When
the magnetospheric radii $R_{\rm m}$ of such stars exceeds their
Keplerian corotation radii $R_{\rm c}$, the effective potential
experienced by charged particles \added{along any given field line
}has a local minimum outside $R_{\rm c}$, which enables charged
material to accumulate in a torus.  Warps can occur when there is
misalignment between the spin and magnetic axes.  In general,
the\added{ regions} in these centrifugally-supported
magnetospheres\added{ with the densest plasma accumulations} need
neither transit nor be opaque in broadband optical light.

@@

To summarize, the lines of evidence in favor of warped plasma tori for
TIC\,141146667 and CPVs more generally are that {\it i)}
Figure~\ref{fig:spec} shows two emitting plasma clumps separated
180$^\circ$ in phase, with double-peaked H$\alpha$ morphology at
quadrature; {\it ii)} the ``W'' shaped photometric eclipses are
predictions of such a model; {\it iii)} CPV dips last 5-50\% of each
cycle, suggesting an extended azimuthal distribution of circumstellar
material.

@@

Microphysically, it is not obvious whether hydrogen alone can produce
the broadband chromatic flux variations seen in CPVs.  The
alternative idea is that charged dust could provide most of the
opacity \citep{Sanderson2023}.  For
TIC\,141146667, Figures~\ref{fig:fulllc}
and~\ref{fig:spec} show that the transits of the H$\alpha$ emitting
material only partially overlap the complex photometric modulation.
This implies that some of the complex photometric dips must involve
additional opacity sources or spatially distinct structures.

@@

The other analog for the CPVs are the $\sigma$~Ori~E variables, a rare
subset of B stars with radiatively-driven winds which accumulate into
warped plasma tori \citep{Townsend2005}.  These tori tend
to have dense antipodal accumulations of plasma sculpted by
tilted-dipole magnetic fields, and the transits of these clumps produce
broadband photometric variability through bound-free scattering
\citep{Townsend2005} and Thomson scattering \citep{Berry2022}.  For
$\sigma$~Ori~E and almost all of its analogs, the result is simple
light curves that resemble those of eclipsing binaries, and time-dynamic
H$\alpha$ spectra similar to those in Figure~\ref{fig:spec}.  The two
known exceptions, HD~37776 and HD~64740, show complex light curves
resembling CPVs \citep{Mikulasek2020,Bouma2024} and have
spectropolarimetric magnetic field maps indicating strong contributions
from higher-order magnetic moments \citep{Kochukhov2011,Shultz2018}.
This suggests that the photometric complexity of CPVs may be related to
magnetic fields with highly multipolar contributions.

@@

Pressing issues include determining the composition and origin of the
circumstellar material, understanding the exact role of the stellar
magnetic field, and exploring the implied space weather experienced by
the close-in rocky exoplanets that, statistically \citep{Dressing2015},
are likely to be present in most CPV systems.

The material's composition -- either pure plasma or dusty plasma --
can be clarified by time-series infrared spectrophotometry.  While
observations of CPVs in the optical suggest chromaticity consistent
with dust \citep{Tanimoto2020,Gunther2022,Koen2023}, electron
scattering in a plasma transiting over a spotted background star might
also produce chromatic features \citep{Rackham2018}.  \added{This
degeneracy is alleviated in the infrared, where the} composition and
size distribution of any dust that is present could be determined by
measuring the extinction curve for a sample of CPVs from 1-20\,$\mu$m.
Composition and size distributions similar to debris from rocky bodies
seen around white dwarfs \citep{Reach2009} would suggest an extrinsic
origin channel.  Compositions and sizes similar to the interstellar
medium would suggest that dust can condense out from M dwarf winds,
similar to processes that occur around evolved stars
\citep{Marigo2008}.

\added{Future models could take a range of steps to clarify the origin
of CPVs.  A first step for magnetohydrodynamic models that can
currently explain the H$\alpha$ features \citep[e.g.][]{Waugh2022}
would be to incorporate an opacity source that can also reproduce the
photometric dips.  It is likely that this effort will require
exploring what stellar magnetic field topologies reproduce the
observed light curve shapes \citep[e.g.][]{Bouma2024}. While previous
models in the rigid-field approximation have explored both dipolar
\citep{Townsend2008} and multipolar fields \citep{Krticka2022}, dynamo
simulations of fully-convective M dwarfs have suggested that
global-scale mean fields might be confined to a single hemisphere
\citep{Brown2020}; such fields would yield accumulation surfaces quite
different from those that have been explored.  One related challenge
is that CPVs currently suffer a dearth of stellar surface field
measurements, unlike benchmark systems such as AB Dor and V374~Peg
\citep{Jardine2002,Vidotto2011}. This might be alleviated through
future spectropolarimetric observations, or from radio observations
sensitive to the magnetic field structure around the star
\citep[see][]{Callingham2021,Kaur2024}.  Finally, the question of what
microphysical process yields the opaque material also merits future
theoretical attention. While recent advances can produce the
catastrophic cooling that leads to dense, cool clumps embedded in a
hot corona \citep{Daley-Yates2023,Daley-Yates2024}, a self-consistent
treatment of radiation propagating through such clumps remains to be
performed. The question of whether dust grains can nucleate in such
clumps, or whether they need to be externally sourced, deserves
careful attention.}

>>>>

Stars with masses below about half that of the Sun, M dwarfs, are the
only type of star to offer near-term prospects for detecting the
atmospheres of rocky exoplanets with surface water.  Community
investment with JWST is proceeding accordingly
\citep[][]{Redfield2024,TRAPPIST1JWSTCommunityInitiative2024}.  The
question of how M dwarfs influence their planets---especially the
retention of their atmospheres---has correspondingly grown in
importance.  Previous work has established
that most M dwarfs host close-in planets \citep{Dressing2015} that on
average are subject to long disk lifetimes
\citep{Ribas2015}, intense UV radiation \citep{France2016},
and frequent stellar flares 
\citep{Feinstein2020}.  However, despite extensive work in these
areas, the plasma and magnetospheric environments that bathe young,
close-in exoplanets remain challenging to quantify.  Understanding
these environments is crucial because they directly impact atmospheric
retention and\added{, ultimately,} habitability.

@@

The two leading hypotheses for explaining CPVs are either that
transiting clumps of circumstellar material corotate with the star
\citep{Stauffer2017}, or that these stars
represent an extreme in naturally-occurring distributions of starspots
or faculae \citep{Koen2021}.  The main argument against a
starspot-only explanation invokes the timescales and amplitudes of the
sharpest photometric features.  However, no independent evidence has
yet been acquired for the presence of any circumstellar material.  If
such material exists, then the geometric correction from the transit
probability would imply an intrinsic occurrence rate at least a few
times larger than the observed rate, suggesting that these clumps
could exist around $\gtrsim$10\% of M dwarfs during their early lives.

@@

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.925\textwidth]{f2.pdf}
  \vspace{-0.2cm}
  \caption{{\bf Evolution of TIC\,141146667 during the Keck/HIRES
  observation (green bar).}  {\bf a,} TESS simple aperture photometry.
  Data gaps were caused by light from the Earth (BTJD 3356-3358.5) and
  Moon (BTJD 3359.5-3360.5).  Raw two minute
  data are in gray; black time-averages to ten minute sampling.  {\bf
  b-c,} Folded TESS light curve before and after spectroscopy.  Black
  now phase-averages to 100 points per
	3.93\ hour cycle.  During BTJD 3352-3356, a state switch occurred near BTJD
  3353, and the dip at $\phi$$\approx$0.8 \added{became less
  pronounced}.  While the \added{dip spanning $\phi$=0.6-1.1} was present both before and after the
  HIRES sequence, its photometric shape evolved during the data gap.
  The orange bar denotes times of spectroscopic transits for the inner
  two H$\alpha$ clumps observed with HIRES (see
  Figure~\ref{fig:spec}).}
  \label{fig:fulllc}
\end{figure*}

\begin{figure*}[!t]
  \centering
  \includegraphics[width=0.99\textwidth]{f3.pdf}
  \vspace{-0.3cm}
  \captionsetup{labelformat=moviefmt,labelsep=colon}
	\caption{\textbf{Emission from circumstellar plasma orbiting
TIC\,141146667.}
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{{\bf
  available here}},
  shows the spectral evolution over five hours.
  {\bf a,} Average TESS light curve from 5 February 2024 to 26
  February 2024 folded on the \periodhr\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf b,} Keck/HIRES H$\alpha$ spectra acquired on 17
  February 2024.  The continuum is set to unity, and the darkest color
  is at twice the continuum to accentuate emission outside the line
  core ($|v/v_{\rm eq}|>1$, for the equatorial velocity $v_{\rm eq}$=130\,\kms).  While
  emission in the line core comes from the star's chromosphere,
  the sinusoidal emission features are most readily described by a
  warped plasma torus.
  {\bf c,} Individual epochs of Panel b, visible in the
  online movie.  The dotted line shows the time-averaged spectrum,
  $f_{\langle t \rangle}$.
  {\bf d,} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf e,} As in Panel b, after subtracting the time-averaged
  spectrum.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch
  is set to mirror the dynamic range of the data.
  {\bf f,} Individual epochs of Panel e, visible in the online
  movie.}
  \label{fig:spec}
\end{figure*}

@@

We observed TIC\,141146667 ($V$=16.2) for five hours on
17~February~2024 using the High Resolution Echelle Spectrometer
(HIRES; \citealt{vogt_hires_1994}) on the 10\,m Keck I telescope.  The
observations spanned the second half of the night, from 10:47 to 16:13
(UT). 
%The star's airmass spanned $z$=1.2-2.2, and we opted for a fixed 15
%minute cadence, except for a final 10 minute exposure due to
%increasing sky brightness at sunrise. 
\added{We opted for a fixed 15 minute cadence, and} used the C2 decker
(0$\farcs$86$\times$14$\farcs$0) in the red configuration, yielding a
resolution $R$$\approx$45{,}000 ($\delta
v$$\approx$6.7\,\kms).
%We binned the CCD readout by a factor of three in the spatial
%dimension, yielding $\approx$1,000 photons (S/N=33) per pixel in the
%continuum at 6500\,\AA. 
Strong winds contributed to 1$''$-1\farcs5 seeing, but conditions were
otherwise
favorable.  We reduced the echelleogram %to a one-dimensional spectrum
using the standard techniques of the California Planet Survey
\citep{Howard2010}.  
%Figure~\ref{fig:spec} shows the result in the
%vicinity of H$\alpha$ without any additional processing.

TESS observed TIC\,141146667 ($T$=13.3) for six non-contiguous months
spanning 2019-2024.  TESS acquired these observations at two-minute
cadence during Sectors 41, 48, and 75.   The 30-minute data during Sectors
14, 15, and 21 smeared sharp features over the star's 3.93~hour
period (see \citealt{Gunther2022}).  The movie in
Figure~\ref{fig:lc} shows the two-minute data: 
like other CPVs, TIC\,141146667 maintains a fixed period over
timescales of years while its detailed photometric morphology evolves.
The nearest known star, TIC\,141146666 ($T$=14.5), is 25$''$ from
TIC\,141146667 and is photometrically \added{quiet};
crowding is not a concern.

We observed with Keck during Sector~75 in an attempt to obtain
simultaneous observations.   Figure~\ref{fig:fulllc} shows the result;
Earth passed within 25$^\circ$ of the TESS camera's boresight from BTJD
3356.0-3358.5, which caused a data gap that ended twelve hours after our
Keck/HIRES observations (green bar).  From BTJD 3359.4-3362.0, the Moon
then passed within 25$^\circ$ of the camera's boresight.  Based on the
level of scattered light in the optimal aperture
\citep{Jenkins2016}, we manually masked out data in the TESS light curve
from 3359.4-3360.1; the remainder of the data during the lunar
approach were usable.  Small gaps from BTJD 3353.6-3353.8 and
3360.1-3360.3 were caused by data downlinks at the spacecraft's
perigee and apogee.

While the data gap is unfortunate, Figure~\ref{fig:fulllc} shows that
both before and after the HIRES data were acquired, a large flux
decrement spanned roughly half of each cycle.  From BTJD 3352-3356, this
dip had two sharp local minima;  the minimum at $\phi$$\approx$0.8
became less pronounced following the flare at BTJD 3353, yielding a dip more
closely resembling an asymmetric ``V'' than a ``W''.  Similar CPV state
changes have been previously seen \citep{Stauffer2017,Bouma2024}.  The
photometric shape therefore evolved during the twelve cycles spanning
the 3356-3358 gap, since the average shape from 3358-3363 more closely
resembles the initial ``W''.  Nonetheless, the global photometric
morphology---a small brightening over 45\% of the period, followed by a
complex flux dip spanning 55\% of the period---is similar before and
after the data gap.

@@

Figure~\ref{fig:spec} shows the TESS and HIRES data from February
2024.  The spectra \added{(Panels b,c)} show emission in
H$\alpha$ beyond the star's equatorial velocity, $v_{\rm eq}$, of
130\,\kms.  There are at least two distinct emission components,
180$^\circ$ apart in phase.  The inner component at lower
velocities has clearer sinusoidal behaviour in time and is
double-peaked, with peak semi-amplitudes of 2.07\,$v_{\rm eq}$ and
2.88\,$v_{\rm eq}$ (see Appendix~\ref{subsec:halpha}).  There
is significant non-periodic variability in the emissivity of this
double-peaked component: the flux excess begins with an amplitude 70\%
above the continuum, and diminishes to 10\% by sunrise.  The
higher-velocity component 180$^\circ$ opposite in phase is detected
only from $\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this outer component
appears connected to the star in velocity space.  While its peak
semi-amplitude of 3.88\,$v_{\rm eq}$ is achieved at both $\phi$=0.25
and $\phi$=0.75, its amplitude similarly decreases from a 60\% excess 
to a 10\% excess over the observation sequence.  In
Appendix~\ref{subsec:halpha}, we measured the period for all three
emission components to be consistent with the
photometric \periodhr\ hour period, to within two minutes for the
\added{inner} component.

These sinusoidal emission features require circumstellar clumps of
partially-ionized hydrogen to corotate with the star.  Based on the
observed sinusoidal periods and velocities, this \added{material is
not moving on a Keplerian orbit}; it is \added{magnetically forced to
co-rotate with the star}.  The velocity
semi-amplitude of the sinusoids gives the mean distance of each clump
from the stellar \added{rotation axis}: 2.07\,$R_\star$ and
2.88\,$R_\star$ for the inner clumps, and 3.88\,$R_\star$ for the
outer clump.   These clumps transit in front of the star when passing
from negative to positive velocity.  The transits of the two inner
clumps last $\approx$22\% of each cycle, from $\phi$=-0.1 to
$\phi$=+0.12.  This spectroscopic
transit coincides with the latter half of the complex eclipse feature
in the TESS data \added{(see Figure~\ref{fig:fulllc} and Panels a and
b of Figure~\ref{fig:spec})}.

The H$\alpha$ line core\added{, visible in Panels e and f of
Figure~\ref{fig:spec}, }is more complex.  At $|\Delta v / v_{\rm
eq}|<1$, most observed H$\alpha$ photons come from the star's
chromosphere; \added{variability in the stellar emission line can be caused
by circumstellar material, bright regions, or dark regions
crossing the star's surface}. 
Figure~\ref{fig:spec}e suggests that \added{all three effects occur in
TIC\,141146667}.
For instance, from $\phi$=0-0.3, double-peaked emission is visible
both on and off-limb; this feature is circumstellar in origin.
However, the emission feature that crosses the star from
$\phi$=0.4-0.9 is brighter than the circumstellar clumps, and its slow
speed instead suggests it is associated with bright region on the
star's surface.  Similarly, from $\phi$=0.6-1.15 a 20\% deep
absorption feature slowly crosses the H$\alpha$ line profile.  This
feature could be either a chromospherically dark region (e.g.~a spot
group), or an azimuthally extended component of the circumstellar
material.  The origin of the other bright and dark streaks passing
across the line core \added{in Figure~\ref{fig:spec}e }are similarly
ambiguous.

\added{Figure~\ref{fig:spec}d shows} a final exercise to
quantify the behavior of the line core, 
by summing the \added{H$\alpha$} flux at $|\Delta v / v_{\rm
eq}|<1$.  This panel shows that changes in the line core flux
\added{($f_{\rm H\alpha\ core}$)} correlate with the broadband
variability throughout most of the light curve, except near
$\phi$$\approx$0.5, corresponding to the transit of the 3.9\,$R_\star$
clump and the occultation of the lower-velocity clump.
\added{Spectroscopic transits may therefore not always be associated
with photometric transits.}

@@

While the H$\alpha$ emission seems to mostly originate from clumps,
the complex flux dip in Figure~\ref{fig:spec} spans roughly half of
each cycle; in CPVs more broadly, individual dips can last 5-50\% of
each cycle \citep{Bouma2024}.  This implies that
azimuthally-distributed material may be the norm for CPVs; a warped
torus may be a more accurate picture than a clump.

\citet{Townsend2005} outlined the physics of how rapidly-rotating
stars with strong magnetic fields can support plasma tori \added{\citep[see
also][]{Nakajima1985,Ferreira2000,Petit2013,Daley-Yates2024}}.  When
the magnetospheric radii $R_{\rm m}$ of such stars exceeds their
Keplerian corotation radii $R_{\rm c}$, the effective potential\added{
  (gravitational plus centrifugal)} experienced by \added{plasma
along any given field line }has a local minimum outside $R_{\rm c}$,
which enables charged material to accumulate in a torus.  Warps can
occur when there is misalignment between the spin and magnetic axes.
In general, the\added{ regions} in these centrifugally-supported
magnetospheres\added{ with the densest plasma accumulations} need
neither transit nor be opaque in broadband optical light.

@@

To summarize, the evidence for warped plasma tori
in TIC\,141146667 and CPVs more broadly is that {\it i)}
Figure~\ref{fig:spec} shows two emitting plasma clumps separated
180$^\circ$ in phase, with double-peaked H$\alpha$ morphology at
quadrature; {\it ii)} ``W'' shaped photometric eclipses are
predictions of such a model; {\it iii)} CPV dips last 5-50\% of each
cycle, suggesting an azimuthally-extended distribution of circumstellar
material.  \added{Two key challenges however are that the warped torus
model predicts two W-shaped eclipses per cycle, and it also predicts
that the antipodal clumps should be equidistant from the star.
Plausible resolutions could invoke different opacities for each clump, 
and non-dipolar stellar magnetic fields.
Generally, the evidence seems to support a warped torus as a
reasonable first approximation, which we hope will be refined by
future efforts.}

@@

Microphysically, it is not obvious whether hydrogen alone can produce
the broadband chromatic flux variations seen in CPVs.  The
alternative is that charged dust could provide most of the
opacity \citep{Sanderson2023}.  For
TIC\,141146667, Figures~\ref{fig:fulllc}
and~\ref{fig:spec} show that the transits of the 
inner H$\alpha$ clump only partially overlap the complex
photometric modulation.  This implies that some of the complex
photometric dips must involve additional opacity sources or spatially
distinct structures.

@@

The other analog for the CPVs are the $\sigma$~Ori~E variables, a rare
subset of B stars with radiatively-driven winds which accumulate into
warped plasma tori \citep{Townsend2005}.  These tori tend
to have dense antipodal accumulations of plasma sculpted by
tilted-dipole magnetic fields, and the transits of these clumps produce
broadband photometric variability through bound-free scattering
\citep{Townsend2005} and Thomson scattering \citep{Berry2022}.  For
$\sigma$~Ori~E and almost all of its analogs, the result is simple
light curves that resemble those of eclipsing binaries, and time-dynamic
H$\alpha$ spectra similar to those in Figure~\ref{fig:spec}.  The two
known exceptions, HD~37776 and HD~64740, 
have spectropolarimetric magnetic field maps indicating strong
contributions from higher-order magnetic moments
\citep{Kochukhov2011,Shultz2018} and complex light curves
resembling CPVs \citep{Mikulasek2020,Bouma2024}.
This suggests that the photometric complexity of CPVs may be related to
magnetic fields with highly multipolar contributions.  
\added{While this suggestion is consistent with the wide range of
axisymmetric and non-axisymmetric topologies seen in mid- and late-M
dwarfs
(e.g.~\citealt{Donati2006,Kochukhov2017,Shulyak2019,Bellotti2024}),
future work is needed to verify whether CPVs indeed select for
complex, non-axisymmetric fields.}

@@

Pressing issues include determining the composition and origin of the
circumstellar material, understanding the exact role of the stellar
magnetic field, and exploring the implied space weather experienced by
the close-in rocky exoplanets that, statistically \citep{Dressing2015},
are likely to be present in many CPV systems.

The material's composition -- either pure plasma or dusty plasma --
can be clarified by time-series infrared spectrophotometry.  While
observations of CPVs in the optical suggest chromaticity consistent
with dust \citep{Tanimoto2020,Gunther2022,Koen2023}, electron
scattering in a plasma transiting over a spotted star might
also produce chromatic features \citep{Rackham2018}.  \added{This
degeneracy is alleviated in the infrared, where the} composition and
size distribution of any dust that is present could be determined by
measuring the extinction curve for a sample of CPVs from 1-20\,$\mu$m.
Composition and size distributions similar to debris from rocky bodies
seen around white dwarfs \citep{Reach2009} would suggest an extrinsic
origin channel.  Compositions and sizes similar to the interstellar
medium would suggest that dust can condense out from M dwarf winds,
similar to processes that occur around evolved stars
\citep{Marigo2008}.

\added{Future models could take a range of steps to clarify the origin
of CPVs.  A first step for magnetohydrodynamic models that can
currently explain the H$\alpha$ features \citep[e.g.][]{Waugh2022}
would be to incorporate an opacity source that can reproduce the
photometric dips.  This effort will also require
exploring what stellar magnetic field topologies reproduce the
observed light curve shapes.  While previous
models in the rigid-field approximation have explored both dipolar
\citep{Townsend2008} and multipolar fields \citep{Krticka2022}, dynamo
simulations of fully-convective M dwarfs have suggested that
global-scale mean fields might be confined to a single hemisphere
\citep{Brown2020}; such fields would yield accumulation surfaces 
different from those that have been explored.  One related challenge
is that CPVs currently suffer a dearth of stellar surface field
measurements, unlike benchmark systems such as AB Dor and V374~Peg
\citep{Jardine2002,Vidotto2011}. This might be alleviated by
future spectropolarimetry, or by radio observations
sensitive to the magnetic field structure around the star
\citep[][]{Callingham2021,Kaur2024}.  Finally, the question of what
microphysical process yields the opaque material also merits future
theoretical attention. While recent advances can produce the
catastrophic cooling that leads to dense, cool clumps embedded in a
hot corona \citep{Daley-Yates2023,Daley-Yates2024}, a self-consistent
treatment of radiation propagating through such clumps has yet to be
performed. Whether dust grains can form in these plasma clumps
similarly deserves attention.}
